---
title: "BCS BIO 5504 - Week Two Assignment"
author: "Ben Stalheim"
date: "2026-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
```

# **Data Dictionary:**

This dataset contains detections of bird species using passive acoustic monitoring and a machine learning model called BirdNET. Included in the data are information on date, time, site, location, and time within the raw audio file. The dataset spans 2023-2025 and data were collected from 27 monitoring points (sites) within 3 different locations. All data were collected in southeastern Georgia, USA, during June and July.

-   Load the dataset (I already did quite a bit of cleaning to this dataset). This is just showing the first few rows:

```{r, include=FALSE}
birdnet_data <- read_csv("bn_dat_allyears.csv") |> 
  select(common_name, sp_code, confidence, date, time, year, site, location)
```

```{r, echo = FALSE}
head(birdnet_data)
```

# Data Structure

-   This shows the names of the columns, types of data, and dimensions of the dataset:

```{r}
colnames(birdnet_data)
str(birdnet_data)
dim(birdnet_data)
```

# Data Summary

-   Counts of the number of detections for each species in the dataset:

```{r, echo=FALSE}
raw_detections <- birdnet_data |> 
  count(common_name, name = "detections") |> 
  arrange(desc(detections))
head(raw_detections)
```
